---
title: "Programming and data analysis in R"
author: "Krzysztof Pleskot  
krzypl@amu.edu.pl  
Room 170"
date: "2025-03-07"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Programming


Programming is the process of creating a series of instructions that tell a computer how to perform a specific task  

## Why does the world need programming?

- Automation of tasks
- Ease of replicating and expanding specific solutions
- Reproducibility

## Is programming only for programmers?

## What is essential for programming?

- Time
- Patience
- Enthusiasm
- Basic mathematical knowledge

## Programming Language

- **Definition** – A set of rules that define when a sequence of symbols forms a computer program and what computations it describes (from Wikipedia).
- There is no single best programming language.

## R as a Programming Language

- Open source
- Well suited for data analysis and visualization
- Extendable functionality through additional packages
- Can be used with an Integrated Development Environment (IDE)

**Disadvantages:**
- Contains many inconsistencies

## Where to learn more?

- <https://r4ds.had.co.nz/> - R for Data Science
- <https://jrnold.github.io/r4ds-exercise-solutions/> - Solutions for R4DS
- <https://couttsgeodata.netlify.app/post/2021-02-28-r_geoscience/> - R packages for geological data analysis
- <https://stackoverflow.com/questions/tagged/r>
- Google

## Course Materials

<https://github.com/krzypl/programming_and_data_analysis_in_r>

## A Bit of Inspiration

```{r echo = TRUE, message=FALSE, warning=FALSE, out.width="50%"}
library(GGally)
data(tips, package = "reshape")
ggpairs(tips[, c(1, 3, 4, 2)],
  upper = list(continuous = "density", combo = "box_no_facet"),
  lower = list(continuous = "points", combo = "dot_no_facet"))
```

## Another Inspiration

```{r echo = FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(maps)
library(viridis)
UK <- map_data("world") %>% filter(region=="UK")
data <- world.cities %>% filter(country.etc=="UK")
mybreaks <- c(0.02, 0.04, 0.08, 1, 7)

data %>%
 arrange(pop) %>%
 mutate( name=factor(name, unique(name))) %>%
 ggplot() +
    geom_polygon(data = UK, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
    geom_point( aes(x=long, y=lat, size=pop, color=pop), alpha=0.9) +
    scale_size_continuous(range=c(1,12)) +
    scale_color_viridis(trans="log") +
    theme_void() + ylim(50,59) + coord_map() + theme(legend.position="none")
```

## Script to Generate the Map

::: {.small}

```{r eval=FALSE, echo=TRUE}
library(ggplot2)
library(dplyr)
library(maps)
library(viridis)
UK <- map_data("world") %>% filter(region=="UK")
data <- world.cities %>% filter(country.etc=="UK")
mybreaks <- c(0.02, 0.04, 0.08, 1, 7)

data %>% arrange(pop) %>%
 mutate( name=factor(name, unique(name))) %>%
 ggplot() +
    geom_polygon(data = UK, aes(x=long, y = lat, group = group),
                 fill="grey", alpha=0.3) +
    geom_point( aes(x=long, y=lat, size=pop, color=pop), alpha=0.9) +
    scale_size_continuous(range=c(1,12)) +
    scale_color_viridis(trans="log") +
    theme_void() + ylim(50,59) + coord_map() +
  theme(legend.position="none")
```

::: 
