---
title: Presentation of grain size data from the publication xxx and the method of
  obtaining them using R
author: "Krzysztof Pleskot"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Check out the R Markdown cheat sheet: <https://rmarkdown.rstudio.com/lesson-15.HTML>

## At the beginning, provide the bibliographic data of the publication and include the article abstract

insert abstract here

## Then proceed to data acquisition

- Data can come from any location. We are interested in the mean grain size.  

- It's important that the number of observations (samples for which grain size analysis was performed) does not exceed 150.  

- If the dataset you choose has more observations, reduce the data frame accordingly to not exceed the limit.  

- When generating code, remember to use `echo=TRUE` and `warning=FALSE`  

```{r echo=TRUE, warning=FALSE}
library(pangaear)
pang_search <- pg_search(query = 'grain size', bbox = c(163, -78, 164, -76), count = 40)
pang_raw <- pg_data(doi = pang_search$doi[1])
pang_output <- pang_raw[[2]]$data
print(pang_output)
```

## Data modification

- Write your own function that converts data from micrometers/millimeters to the phi scale, or from phi scale to micrometers/milimiters, depending on the units in the original publication
# Apply this function to calculate values for each observation (each sample)

## Present data in the recalculated units in relation to depth in the core. Provide suitable labes for axis, title, etc.

```{r echo=TRUE}
plot(pang_output$`Depth sed [m]`, pang_output$`Grain size mean [phi]`)
title(main = "Grain size data from Barret (2001)")
```

## Include a description of the presented data with a minimum of 100 words
